### 2017-03-25 thesis/results/results2.tex
finished editing up to the discussion

In this chapter, the underlying biological mechanism of the obesity associated signatures from CR data were investigated.
In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
First, the direction of GT pathway associated genetic signatures were resolved, then the pathway associated metagenes were compared with the obesity associated metagenes, and lastly linear models were constructed based on the pathway metagenes and patient \gls{bmi} to predict the obesity associated metagenes.

@@

Pathway associated genetic signatures from the \citet{Gatza2010a} study were first examined in the GT data to make sure the results from their paper was reproducible.

\subsection{Ranking and normalisation methods}
\label{sub:ranking_and_normalisation_methods}

In the \citet{Gatza2010a} study, their data comprised of samples from different data sets from other studies, \gls{mas}-normalised and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
However, the analyses so far have used \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no significant difference in the ranking approaches used.
Correlation and scatter plots of the different combinations of transformation matrices (derived from either the \gls{rma}- or \gls{mas}-normalised GT data) with the  GT data normalised with the different methods were calculated and plotted.
From these results, it was clear the normalisation methods used had the most significant effect on the resulting metagenes, rather than the transformation matrices themselves (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} normalisation method was used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.
Therefore, the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).

One thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.

@@

Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top right hand corner of the heatmap.
Other groups included \gls{ifna}/\gls{ifny}/\gls{tnfa} pathways, \gls{er}/\gls{pr}/p53 pathways and p63/\gls{her2} pathways.
In addition to these highly correlated groups, \gls{stat3}/\gls{tgfb}/Src/\gls{egfr}/Akt pathways showed little correlation with one another.
Comparing these groups with the results presented by \citet{Gatza2010a} (\cref{sec:result_from_gatza2010a_study}), the identified clusters approximately resembled the pathway groups identified in by Gatza \textit{et al.}, which confirmed that the directions of GT pathway metagenes were consistent with those used by Gatza \textit{et al.}.

@@

To see whether the directionality of the GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the transformation matrices and the groupings of the metagenes were examined.
CR, FM and \gls{nzbc} data were \gls{rma}-normalised and transformation matrices of the pathway genetic signatures (derived from GT data set) were applied to the data sets.
The metagenes were plotted in a heatmap (shown in \cref{app:b}), which showed similar groupings  as seen in \cref{fig:gatza_meta_dir}.
This result confirmed that the pathway metagenes were acting similarly in all the data sets in which the transformation matrices have been applied to.

@@

To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked with fractional ranking.
Transformation matrices for the obesity associated genetic signatures were made in the CR data set and the pathway associated genetic signatures were made in the GT data set.
Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and transformation matrices.
The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures (\cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}).

@@

In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the transformation matrices were generated in those data sets.
The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
However, what was unexpected was the fact that there were some genetic signatures that were highly correlated in all of the data sets, whereas the metagene correlation of other signatures were highly variable across different data sets.
For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}).
The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from the GT data set.
This also showed that the tumour biology may be different for different data sets, likely due to the difference in the patient cohorts that were selected for each of the studies.

In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation for the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
This suggested that the obesity associated genetic signatures were relatively consistent across all data sets, and the transformation matrices for these signatures could be made in any of the data sets.
For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data, but none of the obesity associated genetic signatures were specific to a data set.

@@

In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes positively correlated strongly with the obesity metagenes.
Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
However, there seem to be a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.

FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
The fact that the FM metagene did not cluster with any of the CR metagenes showed that the nature of the FM metagene was different from those metagenes found in the CR data set.
Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
This may have been due to the lack of consistency of the Akt pathway signature in the different data sets (\cref{tab:svd_vs_tm_path}).
Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from the other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.

The results in this section suggested that none of the obesity associated genetic signatures were not positively correlated with to any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
However, some pathway signatures were negatively correlated with the obesity associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.

@@

To confirm that the obesity associated metagenes were not related to any of the biological pathway signatures from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
If any of the pathway metagenes were significant in the linear model, this would provide evidence that the significant pathways in the linear model were related to the obesity metagene.
Since most of the obesity metagenes were found in the CR data set, and only the CR and the \gls{nzbc} data sets had \gls{bmi} information for the patients, \gls{nzbc} data set was used as the training data set to create the linear models for obesity metagene score prediction.

@@

For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and all of the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the CR data set were the \gls{pr} metagene and whether the patients were obese or not (\cref{tab:lm_sig_var}).
In the model that predicted the FM metagene scores, the obesity status of the patients and Myc metagene score were significant, but the \gls{pr} metagene score was not (\cref{sec:summary_of_the_linear_models_in_nzbc_data}).

@@

The predicted metagenes from all of the patient \gls{bmi}-related linear models were not significantly related to the original metagene scores in many of the obesity metagenes (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
On the other hand, all of the predicted metagenes generated from the models that used the pathway metagene scores showed significant association with the original metagene scores (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
However, the $R^2$ values for all of the predicted metagenes against the true metagenes were very low ($R^2$ \textless{} 0.4).
All of the plots in \cref{fig:predict_cris_cris,sec:summary_of_the_linear_models_in_nzbc_data} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
Furthermore, the fact that all of the predictions made with the models that contained the \gls{pr} metagene showed greater correlation with the true metagene values suggested that most of the variations in the Cr (and other) obesity metagenes were being explained by the \gls{pr} metagene alone, and the patient \gls{bmi} contributed very little to the prediction.

\begin{figure}[htpb]
	\centering
	\includegraphics[page=1,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=2,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=3,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=4,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=5,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=6,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=7,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=74,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=71,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=72,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\caption[Comparison of the predicted Cr metagene scores with the Cr metagene score from the \gls{nzbc} data]{ Box plot and scatter plots comparing the Cr metagene score predicted by the different linear models constructed from the \gls{nzbc} data, with the Cr metagene score from the \gls{nzbc} data.
	Only the results from the models used to predict the Cr metagene scores are shown.
	The summary of the statistics for the other obesity metagenes in the \gls{nzbc} data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
	P-values and $R^2$-value are as described in previous figures.
	}
	\label{fig:predict_cris_cris}
\end{figure}

@@

Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative assocation of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets.
% An alternative explanation was that the obesity metagenes  were not generalisable in different data sets, and thus the models were predicting the obesity metagenes relatively well in \gls{nzbc} data, but not in the CR data.

\begin{figure}[htpb]
	\centering
	\includegraphics[page=1,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=2,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=3,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=4,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=5,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=6,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=7,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=74,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=71,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=72,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\caption[Comparison of the predicted Cr metagene scores with the Cr metagene score from the CR data]{ Box plot and scatter plots comparing the Cr metagene score predicted by the different linear models constructed from the \gls{nzbc} data, with the Cr metagene score from the CR data.
	Only the results from the models used to predict the Cr metagene scores are shown.
	The summary of the statistics for the other obesity metagenes in the CR data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
	P-values and $R^2$-value are as described in previous figures.
	}
	\label{fig:predict_cr_cris}
\end{figure}

@@

Lastly, since only the six most ``consistent'' pathway metagenes were picked from a total of eighteen pathway metagene scores, linear models were constructed with all eighteen pathway metagenes in a stepwise fashion.
Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contains only the variables that ``matter'' to the final linear model \cref{sec:obesity_metagene_prediction_with_pathway_metagenes}.
Patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) were used to generate the stepwise linear model for each of the obesity metagenes.
The linear models that resulted from the stepwise selection of the variables were summarised in \cref{tab:stepwise_model}.

\begin{longtable}{llr{\bfseries}S}
	\caption[Description of the stepwise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes]{Description of the stepwise linear models constructed from the \gls{nzbc} data to predict all of the obesity metagenes}
	\label{tab:stepwise_model}\\
		Obesity metagene & Variables retained in the model & Estimate & {P-value}\\
	\endfirsthead
	\multicolumn{4}{c}{\tablename\ \thetable{}\ (continued)}\\
	\endhead
		Cr         & Akt         & -0.3415   & \bfseries \num{4.43d-8}\\
				   & \gls{egfr}  & 0.3385   & \bfseries \num{1.94d-05}\\
				   & \gls{er}    & -0.2024  & \bfseries 0.004190\\
				   & \gls{ifny}  & 0.1223   & 0.08700\\
				   & \gls{pr}    & -0.3392  & \bfseries \num{2.09d-5}\\
				   & Ras         & -0.2261  & \bfseries 0.001930\\
				   & Src         & -0.1145  & 0.08177\\
				   & \gls{tgfb}  & 0.5564   & \bfseries \textless{} \num{2d-16}\\
				   & \gls{tnfa}  & -0.2550  & \bfseries 0.006160\\
		CrOl       & Overweight  & 0.04092  & 0.07487\\
				   & Obese       & -0.006381 & 0.7561\\
				   & Akt         & -0.3567 & \bfseries \num{1.89d-07}\\
				   & E2F1        & 0.1602  & \bfseries 0.005411\\
				   & \gls{egfr}  & 0.1486  & 0.06268\\
				   & p53         & 0.3366  & \bfseries \num{4.35d-05}\\
				   & \gls{pr}    & -0.2254 & \bfseries 0.0006390\\
				   & Src         & -0.4288 & \bfseries \num{5.85d-10}\\
				   & \gls{stat3} & -0.1241 & \bfseries 0.009658\\

>>>>

In this chapter, the underlying biological mechanism of the obesity associated signatures from the CR data were investigated.
In doing so, the pathway genetic signatures from the \citet{Gatza2010a} study (GT) were utilised to determine which biological pathways the obesity associated genetic signatures were most similar to.
First, the direction of GT pathway associated genetic signatures were resolved, then the pathway associated metagenes were compared with the obesity associated metagenes, and lastly linear models were constructed based on the pathway metagenes and patient \gls{bmi} to predict the obesity associated metagenes.

@@

\subsection{Ranking and normalisation methods for pathway metagenes}
\label{sub:ranking_and_normalisation_methods}

In the \citet{Gatza2010a} study, their data comprised of samples from different studies which were aggregated into a single data set that was \gls{mas} normalised, and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
However, the analyses so far have used \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no significant difference in the ranking approaches used.

Scatter plots were plotted and the correlations were calculated for each of the GT pathway metagenes derived from either \gls{rma} or \gls{mas} normalised GT data (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
The GT pathway metagenes were generated by using the transformation matrices that were derived from either the \gls{rma} or \gls{mas} normalised GT data.
From these results, it was clear the normalisation methods used on the data set had the most significant effect on the resulting metagenes, rather than the normalisation methods used to generate the transformation matrices (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} was chosen as the normalisation method to be used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.

Another thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
Taken together, it was decided that the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).

@@

Before GT pathway metagenes were compared with the obesity associated metagenes, the direction of the GT pathway metagenes had to be checked to make sure that the metagenes were in the ``correct'' direction (\cref{sub:metagene_direction}).
The correlation of all the pathway metagenes with one another were plotted as a heatmap in \cref{fig:gatza_meta_dir}.
The most prominent group had five pathways (E2F1, \gls{pi3k}, Myc, \gls{bcat} and Ras) that clustered at the top right hand corner of the heatmap.
Other groups included \gls{ifna}/\gls{ifny}/\gls{tnfa} pathways, \gls{er}/\gls{pr}/p53 pathways and p63/\gls{her2} pathways.
In addition to these highly correlated groups, \gls{stat3}/\gls{tgfb}/Src/\gls{egfr}/Akt pathways showed little correlation with one another.
Comparing these groups with the results presented by \citet{Gatza2010a} (\cref{sec:result_from_gatza2010a_study}), the identified clusters approximately resembled the pathway groups identified in by Gatza \textit{et al.}, which confirmed that the directions of GT pathway metagenes were consistent with those used by Gatza \textit{et al.}.

@@

To see whether the directionality of the GT pathway metagenes were being transferred across to other data sets properly, pathway metagenes were generated in other data sets with the same transformation matrices and the groupings of the metagenes were examined.
CR, FM and \gls{nzbc} data were \gls{rma} normalised and transformation matrices of the pathway genetic signatures (derived from the GT data set) were applied to the data sets.
The metagenes were plotted in a heatmap (shown in \cref{sec:correlation_of_the_gt_pathway_metagenes_in_other_cancer_data_sets}), which showed similar groupings  as seen in \cref{fig:gatza_meta_dir}.
This result confirmed that the pathway metagenes were acting similarly in all of the data sets in which the transformation matrices have been applied to.

@@

To decide which data set the transformation matrices for obesity and pathway associated genetic signatures should be generated in, the \gls{svd}- and transformation matrix(TM)-generated metagene scores were compared in all of the data sets.
As in \cref{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}, all data sets were normalised with the \gls{rma} method and metagenes were ranked with fractional ranking.
TMs for the obesity associated genetic signatures were made in the CR data set and the TMs for pathway associated genetic signatures were made in the GT data set.
Metagenes for all of the obesity and pathway associated genetic signatures were generated in all of the data sets with \gls{svd} and TMs.
The Spearman correlation of the \gls{svd}-generated and TM-generated metagene scores were calculated for each genetic signatures (\cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}).

@@

In \cref{tab:svd_vs_tm_path,tab:svd_vs_tm_obs}, the obesity and pathway associated genetic signatures had a correlation of 1 in CR and GT data sets respectively, as the obesity and pathway TMs were generated in the CR and GT data sets, respectively.
The correlations of the obesity and pathway metagenes were variable across different data sets, which was as expected since all the data sets were different from one another.
However, what was unexpected was the fact that there were some pathway genetic signatures that were highly correlated in all of the data sets, and there were others that varied considerably across different data sets.
For example, the \gls{bcat} pathway metagene was highly correlated in all of the data sets (\textgreater{} 0.98), whereas the \gls{stat3} pathway metagene was variable across different data sets, ranging from 0.1902 to 0.9159 (\cref{tab:svd_vs_tm_path}; \cref{sec:comparison_of_the_correlation_of_svd_and_tm_generated_pathway_metagenes}).
The fact that some pathway associated metagenes were consistent across different data sets suggested that some of these genetic signatures were reliable and did not depend on the data set the transformation matrices were derived from.
On the other hand, the genetic signatures that were not consistent across the data sets were likely to be dependent on the data set in which they were derived from, and therefore the transformation matrices for these signatures must be derived from the GT data set.
This also showed that the tumour biology may be different across the data sets, likely due to the difference in the patient cohorts that were selected for each of the studies.

In contrast to the pathway associated genetic signatures, all of the obesity associated genetic signatures showed high correlation for the \gls{svd}- and transformation matrix-generated metagenes across the different data sets.
This suggested that the obesity associated genetic signatures were relatively consistent across all data sets, and the transformation matrices for these signatures could be made in any of the data sets.
For simplicity, it was decided that the transformation matrices for all the genetic  signatures would be made in the GT data set, since there were some pathway associated signatures that were specific to the GT data.

@@

In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes had  a strong positive correlation with the obesity metagenes.
Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
With that said, these obesity associated metagenes seem to have a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.

FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captures by the FM metagene was different from those metagenes found in the CR data set.
Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
This may have been due to the lack of consistency of the Akt pathway signature in the different data sets (\cref{tab:svd_vs_tm_path}).
Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from the other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.

The results in this section suggested that none of the obesity associated genetic signatures were positively correlated with any of the pathway associated genetic signatures from the \citet{Gatza2010a} study.
However, some pathway signatures were negatively correlated with the obesity associated genetic signatures, which meant that the obesity signatures were likely capturing the signals from these signatures  rather than the biological signals from the obesity phenotype of the patient.

@@

To confirm the biological relationship between the obesity associated metagenes with the pathway associated metagenes from the \citet{Gatza2010a} study, linear models were created to predict the obesity metagene scores with the pathway metagene scores.
If any of the pathway metagenes were significant in the linear model, this would provide evidence that the significant pathways in the linear model were related to the obesity metagene.
Since most of the obesity metagenes were found in the CR data set, and only the CR and the \gls{nzbc} data sets had \gls{bmi} information for the patients, \gls{nzbc} data set was used as the training data set to create the linear models for obesity metagene score prediction.

@@

For each of the obesity metagenes identified, seven linear models were created: \gls{bmi}-only, \gls{bmi} status-only, \gls{bmi} and \gls{bmi} status, pathway metagenes-only, and the combinations of pathway metagenes with \gls{bmi} and/or \gls{bmi} status.
The only variables that showed significance in any of the models that predicted the obesity metagenes taken from the CR data set were the \gls{pr} metagene and whether the patients were obese or not (\cref{tab:lm_sig_var}).
In the model that predicted the FM metagene scores, the obesity status of the patients and Myc metagene score were significant, but the \gls{pr} metagene score was not (\cref{sec:summary_of_the_linear_models_in_nzbc_data}).

@@

The predicted metagenes from all of the patient \gls{bmi}-related linear models were not significantly related to the original metagene scores in many of the obesity metagenes (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
On the other hand, all of the predicted metagenes generated from the models that used the pathway metagene scores showed significant association with the original metagene scores (\cref{fig:predict_cris_cris}; \cref{sec:summary_of_the_linear_models_in_nzbc_data}).
However, the $R^2$ values for all of the predicted metagenes against the original metagenes were very low ($R^2$ \textless{} 0.4).
All of the plots in \cref{fig:predict_cris_cris,sec:summary_of_the_linear_models_in_nzbc_data} showed clearly that the data points in the scatter plots were highly variable, suggesting that even though the association between the predicted and the true metagene scores were statistically significant, the variables in the linear models may not be strongly associated to the obesity metagene.
Furthermore, the fact that all of the predictions made with the models that contained the \gls{pr} metagene showed greater correlation with the true metagene values suggested that most of the variations in the Cr (and other) obesity metagenes were being explained by the \gls{pr} metagene alone, and the patient \gls{bmi} contributed very little to the prediction.

\begin{figure}[htpb]
	\centering
	\includegraphics[page=1,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=2,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=3,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=4,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=5,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=6,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=7,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=74,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=71,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=72,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cris_with_cris(cris_model)}
	\caption[Comparison of the predicted Cr metagene scores with the Cr metagene score from the \gls{nzbc} data]{ Box plot and scatter plots comparing the Cr metagene score predicted by the different linear models constructed from the \gls{nzbc} data, with the Cr metagene score from the \gls{nzbc} data.
	Only the results from the models used to predict the Cr metagene scores are shown.
	The summary of the statistics for the other obesity metagene predictions from the \gls{nzbc} data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
	P-values and $R^2$-value are as described in previous figures.
	}
	\label{fig:predict_cris_cris}
\end{figure}

@@

Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative assocation of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets, and overfitting of the model in the \gls{nzbc} data was the likely cause of this negative correlation.

\begin{figure}[htpb]
	\centering
	\includegraphics[page=1,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=2,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=3,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=4,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=5,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=6,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=7,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=74,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=71,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=72,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\includegraphics[page=73,width=0.32\linewidth]{results2/prediction_cr_with_cris}
	\caption[Comparison of the predicted Cr metagene scores with the Cr metagene score from the CR data]{ Box plot and scatter plots comparing the Cr metagene score predicted by the different linear models constructed from the \gls{nzbc} data, with the Cr metagene score from the CR data.
	Only the results from the models used to predict the Cr metagene scores are shown.
	The summary of the statistics for the other obesity metagene predictions from the CR data are shown in \cref{sec:summary_statistics_of_the_predicted_obesity_metagenes_with_sample_bmi_bmi_status_in_nzbc_and_cr_data}.
	P-values and $R^2$-value are as described in previous figures.
	}
	\label{fig:predict_cr_cris}
\end{figure}

@@

Lastly, since only the six most ``consistent'' pathway metagenes were picked from a total of eighteen pathway metagene scores, linear models were constructed with all eighteen pathway metagenes in a stepwise fashion.
Step wise construction of a linear model is where the variables for a model are selected computationally depending on the contribution of the variable to the resulting linear model, and thus contain only the variables that ``matter'' to the final linear model \cref{sec:obesity_metagene_prediction_with_pathway_metagenes}.
Patient \gls{bmi}, \gls{bmi} status and all of the pathway metagenes (Akt, \gls{bcat}, E2F1, \gls{egfr}, \gls{er}, \gls{her2}, \gls{ifna}, \gls{ifny}, Myc, p53, p63, \gls{pi3k}, \gls{pr}, Ras, \gls{stat3}, Src, \gls{tgfb} and \gls{tnfa}) were used to generate the stepwise linear model for each of the obesity metagenes.
The linear models that resulted from the stepwise selection of the variables were summarised in \cref{tab:stepwise_model}.

These models were used to predict the obesity metagenes in the \gls{nzbc} data to test the performance of the models.
All of the models were able to make statistically significant prediction of the original obesity metagenes in the \gls{nzbc} data set (\cref{fig:stepwise_cris}).
Unlike the previous models, the stepwise models showed very high $R^2$ values, where most predictions had $R^2$ \textgreater{} 0.80.
The statistical significance and the strong association of the predicted metagenes were observed in the CR data set as well (\cref{fig:stepwise_cr}).
Though the $R^2$ values were slightly lower than those from the \gls{nzbc} data set, the $R^2$ values from the CR data predictions were more promising than the predictions made by the non-stepwise linear models.
